Curve Average \\ {\large Project 3 for class CS6491 Computer Graphics}

### 2015-11-02 paper/paper.tex
Done with the text. Next: Feedback

\newpage
\section{Average Curve Computation}\label{chapter:Tracing}

@@

\subsection{Distance functions}
TODO
\blindtext

\subsubsection{Newton-Raphson Method}
TODO
\blindtext

\subsubsection{Laguerre's Method}
TODO
\blindtext
\SetArgSty{textup}
\begin{algorithm}[ht!]
  \SetCommentSty{emph}
  \KwIn{$\{P_0, T_0, P_1, T_1\}$: control points, $Q$: input point}
  \KwResult{$t$: time in $[0,1]$ for the closest point}
  $[a,b,c,d,e,f] \leftarrow $generateDistanceCoeffs$(P_0,T_0,P_1,T_1,Q)$; \\
  $res \leftarrow $solveQuintic($a,b,c,d,e,f$); \\
  $\{minDist, t^*\} \leftarrow \{\infty,-1\}$;\\
  \ForEach{$r \in res$} {
    \lIf{$|r.im| > 0$} \textbf{continue}; \\
    \lIf{$r.re < 0\;$ \textbf{or} $\;r.re > 1$} \textbf{continue}; \\
    $P_t \leftarrow \tau(P_0,T_0,P_1,T_1,r.re)$; \\
    \If{$|P_t - Q| < minDist$} {
      $\{minDist, t^*\} \leftarrow \{|P_t - Q|, r.re\}$;\\
    }
  }
  \Return{$t^*$};
  \caption{closestPointOnCubicHermite()}
\end{algorithm}

\subsection{Curve tracing}
TODO
\blindtext

\begin{algorithm}[ht!]
  \SetCommentSty{emph}
  \KwIn{$\{\mathbf{A},\mathbf{B}\}$: curves}
  \KwResult{$\mathbf{C}$: average curve}
  $\mathbf{C}.$add($\mathbf{A}[0], 0$); \\
  $\{current, t_a, t_b, t_a^\prime, t_b^\prime\} \leftarrow \{\mathbf{C}[0],0,0,0,0\}$; \\
  \While{true} {
  $\vec{T}_A \leftarrow \vec{n}$($\mathbf{A}.\tau$($min(1,t_a+\epsilon)$) - $\mathbf{A}.\tau$($min(1,t_a)$)); \\
  $\vec{T}_B \leftarrow \vec{n}$($\mathbf{B}.\tau$($min(1,t_b+\epsilon)$) - $\mathbf{B}.\tau$($min(1,t_b)$)); \\
  $current \stackrel{+}\leftarrow \text{STEP\_SIZE} \;\vec{n}(\vec{T}_A + \vec{T}_B)$; \\
  $\{t_a^\prime, t_b^\prime, r, P_c\} \leftarrow \text{snap}(\mathbf{A},\mathbf{B},current)$; \\
  \lIf{$max(t_a,t_b) < 1$ \textbf{or} $t_a^\prime < t_a$ \textbf{or} $t_b^\prime < t_b$} \textbf{break}; \\
  $\mathbf{C}.$add($P_c$); \\
  }
  \Return{$\mathbf{C}$};
  \caption{traceCurve()}
\end{algorithm}

TODO
\blindtext

@@

\subsection{Average Distance (geodesic) Sampling}
TODO

\subsection{Detect non-compatible control curves}
TODO

@@

% EXAMPLES
\begin{figure*}
	\centering
		\includegraphics[scale=0.4]{images/Result1.png}
		\includegraphics[scale=0.4]{images/Result2.png}
	\caption{Example 1: a bow}
	\label{fig:Result1}
\end{figure*}
\begin{figure*}
	\centering
		\includegraphics[scale=0.4]{images/Result3.png}
	\caption{Example 2: a waterslide}
	\label{fig:Result3}
\end{figure*}
\begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/Result4.png}
		\includegraphics[scale=0.5]{images/Result5.png}
	\caption{Example 3: extreme cases and non-compatibility detection}
	\label{fig:Result4}
\end{figure*}

>>>>

\section{Average Curve Computation}\label{chapter:Tracing}

@@

\subsection{Distance functions}

To satisfy the distance constraints (1) and (2) in the objective function, we need to first 
define the distance functions for a point and a curve. Note that since our curve representation
is composed of multiple splines, we are actually interested in distance between a point
and a hermite (quadratic or cubic). If $P(t)$ represents a point on the spline, parameterized
with time $t$, we are interested in finding $t$ such that the tangent at that point $P^\prime(t)$ and
the vector from that point to the input point $Q$ are perpendicular:

\begin{equation}
 P^\prime(t) \cdot (Q - P(t)) = 0
\end{equation}

If $P(t)$ is a quadratic polynomial in $t$, then the perpendicularity constraint is a cubic in $t$ and
the points where the constraint are satisfied can be solved in closed form. Otherwise, if $P(t)$ is a 
cubic polynomial in $t$, then the perpendicularity constraint is a quintic in $t$ and its roots
need to determined with a local optimizer.

Once all the roots of the perpendicularity constraint is found, the ones with imaginary coefficients
and the ones with real values that are out of the [0,1] bound need to be eliminated. Moreover, to
choose the closest perpendicular point, an iterative comparison has to be executed as shown in 
Algorithm 1 for a quintic function. Note that the coefficients for the polynomial are obtained
through Matlab's symbolic toolbox.

\SetArgSty{textup}
\begin{algorithm}[ht!]
  \SetCommentSty{emph}
  \KwIn{$\{P_0, T_0, P_1, T_1\}$: control points, $Q$: input point}
  \KwResult{$t$: time in $[0,1]$ for the closest point}
  $[a,b,c,d,e,f] \leftarrow $generateDistanceCoeffs$(P_0,T_0,P_1,T_1,Q)$; \\
  $res \leftarrow $solveQuintic($a,b,c,d,e,f$); \\
  $\{minDist, t^*\} \leftarrow \{\infty,-1\}$;\\
  \ForEach{$r \in res$} {
    \lIf{$|r.im| > 0$} \textbf{continue}; \\
    \lIf{$r.re < 0\;$ \textbf{or} $\;r.re > 1$} \textbf{continue}; \\
    $P_t \leftarrow \tau(P_0,T_0,P_1,T_1,r.re)$; \\
    \If{$|P_t - Q| < minDist$} {
      $\{minDist, t^*\} \leftarrow \{|P_t - Q|, r.re\}$;\\
    }
  }
  \Return{$t^*$};
  \caption{closestPointOnCubicHermite()}
\end{algorithm}

\subsection{Optimization methods}

Here, we would like to make a brief mention of the Java Commons Math library which provides a 
plethora of optimizer and share our experience for future users.

\subsubsection{Newton-Raphson Method}
The Newton-Raphson method is one of the few methods that does not need an initial bracketing (limiting
the solution space) and is also conceptually simple: just follow the derivative at the initial sample
until it hits a zero-crossing and repeat until convergence. However, unfortunately, the library
implementation does not always return the closest solution to the sample and thus, hard to use to 
determine all the roots of the quintic function.

\subsubsection{Laguerre's Method}
The Laguerre's method is dependent on two initial bounds which have different signs when the 
quintic is evaluated. The positive aspect of the implementation of the method is that it is guaranteed to find a solution
if one exists within the bounds and if not, throw an exception. Thus, we discretize the
time domain and search for the roots systematically.

\subsection{Curve tracing}

The block diagram that outlines the main functions in our curve tracing algorithm is presented in Figure 1.
The main idea is to start with the intersection of the two curves as the first sample and continuously project
out in the ``medial-line'' of the projected points' tangents iteratively. The traceCurve function in Algorithm 2
demonstrates the steps involved.

\begin{algorithm}[ht!]
  \SetCommentSty{emph}
  \KwIn{$\{\mathbf{A},\mathbf{B}\}$: curves}
  \KwResult{$\mathbf{C}$: average curve}
  $\mathbf{C}.$add($\mathbf{A}[0], 0$); \\
  $\{current, t_a, t_b, t_a^\prime, t_b^\prime\} \leftarrow \{\mathbf{C}[0],0,0,0,0\}$; \\
  \While{true} {
  $\vec{T}_A \leftarrow \vec{n}$($\mathbf{A}.\tau$($min(1,t_a+\epsilon)$) - $\mathbf{A}.\tau$($min(1,t_a)$)); \\
  $\vec{T}_B \leftarrow \vec{n}$($\mathbf{B}.\tau$($min(1,t_b+\epsilon)$) - $\mathbf{B}.\tau$($min(1,t_b)$)); \\
  $current \stackrel{+}\leftarrow \text{STEP\_SIZE} \;\vec{n}(\vec{T}_A + \vec{T}_B)$; \\
  $\{t_a^\prime, t_b^\prime, r, P_c\} \leftarrow \text{snap}(\mathbf{A},\mathbf{B},current)$; \\
  \lIf{$max(t_a,t_b) < 1$ \textbf{or} $t_a^\prime < t_a$ \textbf{or} $t_b^\prime < t_b$} \textbf{break}; \\
  $\{t_a, t_b\} \leftarrow \{t_a^\prime, t_b^\prime\};$ \\
  $\mathbf{C}.$add($P_c$); \\
  }
  \Return{$\mathbf{C}$};
  \caption{traceCurve()}
\end{algorithm}

For a sample \textit{current}, the algorithm first finds the tangents
at the projected point $\vec{T}_A$ and $\vec{T}_B$, and \textit{projects out} the current sample by their median.
However, since this point is not guaranteed to represent the distance constraints (1) and (2), nor since that
sphere would be minimal, we need to incrementally move it with the \textit{snap} function.

The pseudo-code for the \textit{snap} function is presented in Algorithm 3. At each iteration,
two main changes are made on the initial sample. First, the point is moved towards its closest projections
as shown in the lines 3-7. Then, the point is moved towards the planes defined by the projections 
and the sphere center to minimize the radius of the sphere (in lines 8-10).

@@

\subsection{Average Distance (geodesic) Sampling}

One challenge with the output of the naive tracing algorithm is that the samples on the curve $C$
can have arbitrary relative distances. This can be an issue when additional visualization tools
are built upon this curve such as tubes and boardwalk patterns. One approach is to predefine
the number of points desired and sample the curve $C$ based on the total arc length and the 
chosen resolution.

Algorithm 4 outlines this approach where the relevant $C$ samples for the new sample are
determined with a preliminary search (lines 6-12) and then, an intermediary point is computed
with linear interpolation. Next, the interpolated point is pushed back on to the curve $C$
to satisfy the object constraints 1-3 using the snap function in lines 14-15.

The effect of the average distance sampling in comparison to the original average curve
$C$ can be observed in Figure 3 between items (a) and (b). As expected, the samples
along the geodesic curve is much more refined and can sample more thoroughly the ``incompatible''
regions.

\begin{algorithm}[ht!]
  \SetCommentSty{emph}
  \KwIn{$\{\mathbf{A},\mathbf{B}\}$: curves, $N:$ number of points}
  \KwResult{$\mathbf{G}$: geodesic average curve}
  $\mathbf{C} \leftarrow$ trace($\mathbf{A}, \mathbf{B}$); \\
  $length \leftarrow \mathbf{A}.$arcLength()$ + \mathbf{B}.$arcLength(); \\
  $\{\delta l, i, \mathbf{G}\} \leftarrow \{length / N, 0, \{\mathbf{C}[0]\}\}$;\\
  \While{true} {
    $\{g, \underline{i}, \underline{d}, \overline{i}, \overline{d}\} \leftarrow \{\mathbf{G}$.last, $i,0,-1,0\};$ \\
    \While{$i < \mathbf{C}.$length} {
      $d_A \leftarrow \mathbf{A}.$arcLength($proj(g,\mathbf{A}$), $proj(C[i], \mathbf{A}$)); \\
      $d_B \leftarrow \mathbf{B}.$arcLength($proj(g,\mathbf{B}$), $proj(C[i], \mathbf{B}$)); \\
      \lIf{$d_A + d_B < \delta l$} $\{\underline{i}, \underline{d}\} \leftarrow \{i, d_A + d_B\}$; \\
      \Else {
	$\{\underline{i}, \underline{d}\} \leftarrow \{i, d_A + d_B\};$ \\
	\textbf{break};
      }
    }
    \lIf{$i \geq \mathbf{C}.$length} break; \\
    $P_c \leftarrow $LERP$(\mathbf{C}[\underline{i}], \mathbf{C}[\overline{i}], (\delta l - \underline{d}) / (\overline{d} - \underline{d}))$;\\
    $\mathbf{G}$.add(snap($\mathbf{A}, \mathbf{B}, P_c$)); \\
  }
  \Return{$\mathbf{G}$};
  \caption{geodesicTrace()}
\end{algorithm}

\newpage
\subsection{Detect incompatible control curves}

An important aspect of the average curve formulations provided in the first section is that the 
there may not always be a solution to the three fundamental constraints of the average curve.
In such cases, the average curve may not (1) exist, (2) be smooth, or (3) be close to the medial axis,
and thus, it is important to inform the user when such cases may arrive as they manipulate the control points.

One method of doing so is to monitor the arclength between the consecutive points in the average curve (geodesic)
trace. If the arclength is significantly larger than the expected step size (1.5 times more in our implementation), 
then a warning can be given by tracing the corresponding curve sections in red. Figure 2 below demonstrates
examples of such extreme cases and the incompatible sections of the input control curves that would lead to a branching
in the medial axis.

\begin{figure}[ht!]
  \includegraphics[width=1.0\linewidth]{images/Result45.png}
	\caption{Extreme cases and non-compatibility detection}
	\label{fig:Result4}
\end{figure}

@@

% EXAMPLES
\begin{figure*}
	\centering
		\includegraphics[scale=0.4]{images/Result1.png}
		\includegraphics[scale=0.4]{images/Result2.png}
	\caption{Example 1: a bow}
	\label{fig:Result1}
\end{figure*}
\begin{figure*}
	\centering
		\includegraphics[scale=0.4]{images/Result3.png}
	\caption{Example 2: a waterslide}
	\label{fig:Result3}
\end{figure*}
